---
title: "vis.cap.articles"
author: "Aarsh Batra"
date: "`r Sys.Date()`"
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## load libraries and background helper functions
```{r load_lib}
library(tidyverse)
library(jsonlite)
library(here)
library(sf)
library(ggplot2)
library(geojson)
library(devtools)
library(geojsonio)
library(geojsonsf)
library(roxygen2)
library(magrittr)
library(stringr)
library(data.table)
library(viridis)
library(ggthemes)
library(hrbrthemes)
library(RColorBrewer)
library(haven)
library(janitor)
library(lwgeom)
library(rvest)
library(dplyr)
library(ggrepel)
library(ggsflabel)
library(fuzzyjoin)
library(tidytext)
library(tm)
library(wordcloud2)
library(recipes)
library(forcats)
library(forcats)
library(geofacet)
library(pak)
library(camcorder)
library(ggbeeswarm)
library(ggtext)
library(ggimage)
library(tidyverse)
library(ggbeeswarm)
library(cowplot)
library(ggtext)
library(scales)
library(grid)
library(png)
library(here)
library(ggimage)
library(showtext)

# Negation of the "%in%" operator function
`%notin%` <- Negate(`%in%`)


```

## Story 2: Objects launched in Space, tidy tuesday dataset (Voronoi app submission)

```{r}
#### Read directly from GitHub and clean column names
outer_space_objects <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-04-23/outer_space_objects.csv')

outer_space_objects <- outer_space_objects %>% 
  clean_names() %>%
  mutate_if(is.character, tolower)


#### group data into decades--------------
outer_space_objects <- outer_space_objects %>%
  mutate(decade = paste0(floor(year / 10) * 10, "-", floor(year / 10) * 10 + 10)) %>%
  arrange(year) %>%
  filter(entity != "world")


#### calculate the total number of objects sent in space by each entity in each decade
outer_space_objects_summary <- outer_space_objects %>%
  group_by(entity, decade) %>%
  summarise(num_objects = sum(num_objects, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(entity, decade, num_objects) %>%
  mutate(entity = str_to_title(entity)) 

# #### outer space objects summary with chat GPT filled continents
# outer_space_objects_summary <- readxl::read_xlsx("./data/intermediate/obj_launched_in_outer_space/outer_space_objects_summary_with_continents.xlsx")


#### Add a continent column



#### Plot the number of unique entities that sent objects into space in each decade

# Load rocket image (replace with the path to your rocket image)
rocket_img <- grid::rasterGrob(png::readPNG(paste0(here(), "/images/thin_rocket_bg_rm_adobe.png")), width = unit(1, "npc"), height = unit(1, "npc"))
satellite_img <- png::readPNG(paste0(here(), "/images/tiny_satellite.png"))
background_img <- grid::rasterGrob(readPNG(paste0(here(), "/images/starry_night_2_png.png")), width = unit(1, "npc"), height = unit(1, "npc"))

# Create the plot
plt <- outer_space_objects_summary %>%
  ggplot() +
    # annotation_custom(background_img, ymin = -Inf, ymax = Inf, xmin = -Inf, xmax = Inf) +
  
  # Main plot with geom_quasirandom, using deep space colors
  geom_quasirandom(mapping = aes(x = num_objects, y = decade, size = num_objects, alpha = num_objects),
                   groupOnX = FALSE, width = 0.4, color = "cyan") + 
  scale_size(range = c(2.7, 15), breaks = c(1, 10, 100, 1000, 5000)) +
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) + 
  scale_alpha_continuous(range = c(0.25, 0.8), breaks = c(1, 10, 100, 1000, 5000)) +
  
  # # Highlighting Russia and USA with elegant space-themed colors
#   geom_point(data = filter(outer_space_objects_summary, num_objects > 300),
#              aes(x = num_objects, y = decade, size= num_objects, color = entity)) +
# scale_colour_paletteer_d("dichromat::Categorical_12") +
  
 # Labels for high object counts > 1000
  geom_text(data = filter(outer_space_objects_summary, num_objects > 800),
            aes(x = num_objects, y = decade, label = paste(entity)),
            hjust = 0.5, vjust = -1.2, size = 3.5, color = "white") +
  
  #  # Labels for high object counts
  geom_text(data = filter(outer_space_objects_summary, num_objects > 100, num_objects < 1000, decade %in% c("1950-1960", "1960-1970", "1970-1980", "1980-1990")),
            aes(x = num_objects, y = decade, label = paste(entity)),
            hjust = 0.5, vjust = -1.2, size = 3.5, color = "white") +
  # 
    geom_text(data = filter(outer_space_objects_summary, num_objects < 100, decade %in% c("1950-1960")),
            aes(x = num_objects, y = decade, label = paste(entity)),
            hjust = 0.5, vjust = -1.2, size = 3.5, color = "white") +
  
      geom_text(data = filter(outer_space_objects_summary, num_objects > 100, decade %in% c("2000-2010", "2010-2020", "2020-2030")),
            aes(x = num_objects, y = decade, label = paste(entity)),
            hjust = 0.5, vjust = -1.2, size = 3.5, color = "white") +
  
       geom_text(data = filter(outer_space_objects_summary, decade %in% c("1960-1970"), num_objects > 8),
            aes(x = num_objects, y = decade, label = paste(entity)),
            hjust = 0.5, vjust = -1.2, size = 3.5, color = "white") +
  
 ggplot2::annotate(
    "text", label = "Japan, France",
    x = 23 , y = "1970-1980" , size = 3.5, colour = "white"
  ) +
  
           geom_text(data = filter(outer_space_objects_summary, num_objects > 25, num_objects < 100, decade %in% c("1980-1990")),
            aes(x = num_objects, y = decade, label = paste(entity)),
            hjust = 0.5, vjust = -1.2, size = 3.5, color = "white") +
  
   ggplot2::annotate(
    "text", label = "US, Russia",
    x = 650 , y = "1990-2000" , size = 3.5, colour = "white"
  ) +
  
             geom_text(data = filter(outer_space_objects_summary, num_objects > 60, num_objects <= 150, decade %in% c("2000-2010")),
            aes(x = num_objects, y = decade, label = paste(entity)),
            hjust = 0.5, vjust = -1.2, size = 3.5, color = "white") +
  
             geom_text(data = filter(outer_space_objects_summary, num_objects > 30, num_objects < 100, decade %in% c("1990-2000")),
            aes(x = num_objects, y = decade, label = paste(entity)),
            hjust = 0.5, vjust = -1.2, size = 3.5, color = "white") +
  
  
  
    # Log scale for y-axis
  scale_x_log10(breaks = c(1, 10, 100, 1000, 10000), limits = c(1, 10000)) +
  
     # geom_image(aes(x = num_objects, y = decade, image = paste0(here(), "/images/tiny_satellite.png")),
     #        size = 0.05,  # Adjust size of the satellite image
     #        position = position_quasirandom(width = 0.1)) +
  
  # Apply a dark space-themed background with stars as minor gridlines
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",
        legend.text = element_text(color = "white", family = "Audiowide"),
        plot.background = element_rect(fill = "#001f3f", color = NA),  # Deep space color
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        legend.title = element_blank()) +  # Add space for the rocket image
  
  # Add rocket image along the y-axis (adjust width/height and coordinates)
  annotation_custom(rocket_img,
                    ymin = -0.5, ymax = 1.3, xmin = -Inf, xmax = Inf) +

  # Title and subtitle with space-related theme
  labs(title = "**Number of Objects  launched  into  Space by Decade**",
       subtitle = "A cosmic view  of  space &nbsp; launches &nbsp; across &nbsp; the &nbsp; world",
       y = "Decade", 
       x = "Number of Objects Launched", 
       color = "", 
       caption = expression("Data Source · UNOOSA | Data Processed · Our World in Data | Graphic · github.com/AarshBatra")) +
  
  theme(
    plot.title = element_markdown(family = "Audiowide", size = 24, face = "bold", color = "#FFB300", hjust = 0),
    plot.subtitle = element_markdown(family = "Audiowide", size = 16, color = "white", hjust = 0, 
                                     margin = margin(b = 1, unit = "cm")),
    axis.text = element_text(family = "Audiowide", size = 12, color = "white", face = "bold"),
    axis.title.x = element_text(family = "Audiowide", size = 14, face = "bold", color = "white", margin = margin(b = 1, t = 0.5, unit = "cm")),
    axis.title.y = element_text(family = "Audiowide", size = 14, face = "bold", color = "white", margin = margin(l = 0.5, r = 0.5, unit = "cm")),
    axis.line.x = element_blank(), 
    axis.line.y = element_blank(), 
    legend.box.background = element_rect(fill = NA,  color = "white", size = 0.3),
    plot.caption = element_text(family = "Audiowide", size = 9, hjust = 0.5, color = "white", margin = margin(b = 0.2, unit = "cm"))
  ) +
  coord_flip() 

plt1 <- ggbackground(plt, "./images/starry_night_3_png.png")

ggsave("plot.png", plot = plt1, width = 1000, height = 900, units = "px")


```


## Story 1: Understanding the state of global air quality data, research and funding infrastructure


### Part 1: How much longer can you live if you breathed clean air and how does it compare with other health threats? - Air Quality Life Index

```{r echo=TRUE part1fig1}

# load in the aqli helper script (from the R/ subfolder of the repo) and download all relevant datasets as specified in the readme file of the repo.

# Define the target projection (Equal Earth - EPSG:8857)
equal_earth_proj <- "+proj=eqearth +datum=WGS84 +wktext"

# converting AQLI shapefiles to a Equal Earth Projection
gadm2_aqli_2021_shp <- st_transform(gadm2_aqli_2021_shp, equal_earth_proj)
gadm1_aqli_2021_shp <- st_transform(gadm1_aqli_2021_shp, equal_earth_proj)
gadm0_aqli_2021_shp <- st_transform(gadm0_aqli_2021_shp, equal_earth_proj)

# figuring out the objectids for gadm1 for which population is not available
obid_gadm1_pop_na <- gadm1_aqli_2021 %>% 
  filter(is.na(population)) %>%
  select(objectid_gadm1) %>% 
  unlist() %>%
  as.vector()

# figuring out the objectids for gadm0 for which population is not available
obid_gadm0_pop_na <- gadm0_aqli_2021 %>% 
  filter(is.na(population)) %>%
  select(objectid_gadm0) %>%
  unlist() %>%
  as.vector()


# global shapefile gadm2 map data
aqli_global_gadm2_data <- gadm2_aqli_2021 %>%
  left_join(gadm2_aqli_2021_shp, by = c("objectid_gadm2" = "obidgadm2")) %>%
  add_aqli_color_scale_buckets("lyl", "llpp_who_2021") %>%
  select(-geometry, geometry) %>%
  st_as_sf()

## removing extremely tiny polygons from all three shapefiles to focus on the main map

aqli_global_gadm2_data$area <- st_area(aqli_global_gadm2_data)

# minimum area threshold for the polygons
threshold <- 10000000 # units are m^2

# adding a "area without units" column
aqli_global_gadm2_data <- aqli_global_gadm2_data %>% 
  mutate(area_without_units = as.numeric(str_remove(area, " [m^2]"))) 

aqli_global_gadm2_data <- aqli_global_gadm2_data %>%
  filter(area_without_units > threshold) 

# filter out small island regions from all gadm shapefiles
aqli_global_gadm2_data <- aqli_global_gadm2_data %>%
  filter(country %notin% c("Samoa", "American Samoa", "French Polynesia", "Fiji", 
                           "Tonga", "Niue", "Cook Islands", "Vanuatu", "Solomon Islands", 
                           "Vanuatu", "New Caledonia", "Marshall Islands", "Micronesia", 
                           "French Southern Territories", "Saint Helena, Ascension and Tris")) %>%
  filter(!((country  == "New Zealand") & (name_1 == "Galápagos"))) %>%
  filter(!((country  == "United States") & (name_1 == "Hawaii"))) %>%
  filter(!is.na(population)) 

gadm1_aqli_2021_shp <- gadm1_aqli_2021_shp %>%
  filter(name0 %notin% c("Samoa", "American Samoa", "French Polynesia", "Fiji", 
                           "Tonga", "Niue", "Cook Islands", "Vanuatu", "Solomon Islands", 
                           "Vanuatu", "New Caledonia", "Marshall Islands", "Micronesia", 
                           "French Southern Territories", "Saint Helena, Ascension and Tris")) %>%
  filter(!((name0  == "New Zealand") & (name1 == "Galápagos"))) %>%
  filter(!((name0  == "United States") & (name1 == "Hawaii"))) %>%
  filter(obidgadm1 %notin% obid_gadm1_pop_na) 

gadm0_aqli_2021_shp <- gadm0_aqli_2021_shp %>%
  filter(name0 %notin% c("Samoa", "American Samoa", "French Polynesia", "Fiji", 
                           "Tonga", "Niue", "Cook Islands", "Vanuatu", "Solomon Islands", 
                           "Vanuatu", "New Caledonia", "Marshall Islands", "Micronesia", 
                           "French Southern Territories", "Saint Helena, Ascension and Tris")) %>%
  filter(obidgadm0 %notin% obid_gadm0_pop_na) 


# creating a tibble for text boxes annotation positions and labels in order: LA, Delhi, Ouest(Cameroon), Melbourne (Australia), Lima (Peru), Warsaw (Poland)
tb_annotation <- tibble(
  label_x = c(-139, 64.83, -9, 126.5, -87.72, -29.66),
  label_y = c(18.11, 7.5, -0.54, -49.09, -20.34, 40.04),
  text = c("**Los Angeles (US)**, 0.65 years = 7.8 months",
           "**Delhi (India)**, 11.9 years", 
           "**Ouest (Cameroon)**, 4.5 years", 
            "**Melbourne (Australia)** is in compliance with the WHO PM₂.₅ guideline of 5 µg/m³", 
            "**Lima (Peru)**, 2.3 years", 
            "**Warsaw (Poland)**, 1.2 years"))


# global AQLI map for explained.media website
aqli_global_gadm2_map <- aqli_global_gadm2_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), color = "transparent", lwd = 0) +
   geom_sf(data = gadm1_aqli_2021_shp, color = "transparent", fill = "transparent", lwd = 0) +
 geom_sf(data = gadm0_aqli_2021_shp, color = "aliceblue", fill = "transparent", lwd = 0.3) +
   scale_fill_manual(values = c("0 to < 0.1" = "#ffffff", 
                               "0.1 to < 0.5" = "#ffeda0", 
                               "0.5 to < 1" = "#fed976", 
                               "1 to < 2" = "#feb24c", 
                               "2 to < 3" = "#fd8d3c", 
                               "3 to < 4" = "#fc4e2a", 
                               "4 to < 5" = "#e31a1c", 
                               "5 to < 6" = "#bd0026", 
                               ">= 6" = "#800026")) +
    geom_textbox(data = tb_annotation, aes(x = label_x, y = label_y, label = text),
               colour = "white",
               box.colour = "#022531",
               fill = NA,
               family = "Segoe UI",
               size = 6.5) +
    #geom_curve(x = -118.25, xend = -134, y = 34.09, yend = 23.2, color = "white", arrow = arrow(), curvature = 0.5) +
 annotate("segment", x = -118.25, xend = -134, y = 34.09, yend = 23.2, arrow = arrow(type = "closed", length = unit(0.2, "inches")), colour = "white") + # US
 annotate("segment", x = 77.14, xend = 65.09, y = 28.68, yend = 11.82, arrow = arrow(type = "closed", length = unit(0.2, "inches")), colour = "white") +
 annotate("curve", x = 10.18, xend = -12.28, y = 5.5, yend = -20.96, arrow = arrow(type = "closed", length = unit(0.2, "inches")), colour = "white", curvature  = -0.5) +
 annotate("segment", x = 144.97, xend = 131.18, y = -37.97, yend = -41.01, arrow = arrow(type = "closed", length = unit(0.2, "inches")), colour = "white") +
 annotate("segment", x = -77.03, xend = -82.74, y = -12.02, yend = -16.39, arrow = arrow(type = "closed", length = unit(0.2, "inches")), colour = "white") +
annotate("segment", x = 21.03, xend = -19.7, y = 52.24, yend = 43, arrow = arrow(type = "closed", length = unit(0.2, "inches")), colour = "white") +
      coord_sf(default_crs = sf::st_crs(4326)) +
  ggthemes::theme_map() +
labs(title = "How much longer can you live if you breathed clean air?",
     subtitle = expression("Life expectancy gains, if 2021" ~ PM[2.5] ~ "pollution is reduced to meet the WHO" ~ PM[2.5] ~ "guideline"),
     fill = "Potential gain in life expectancy (years)",
caption = expression("Source · Air Quality Life Index, Annual Report 2023 | Graphic · Aarsh Batra · github.com/AarshBatra | explained.media")) + 
  theme(legend.position = c(0.6, 0.98), 
        legend.background = element_rect(fill = "#022531"), 
        legend.text = element_text(color = "white", face = "bold" , size = 21), 
        plot.background = element_rect(colour = "#022531",
                                       fill = "#022531"), 
        plot.caption = element_text(size = 15, hjust = 0.03, color = "white"), 
        plot.title = element_text(size = 35, family = "Segoe UI", face = "bold", color = "white", hjust = 0.03, 
                                  margin = margin(b = 0.5, unit = "cm")),
        plot.subtitle = element_text(size = 22, family = "Segoe UI", face = "italic", color = "white", hjust = 0.03, 
                                     margin = margin(b = 0.5, unit = "cm")),
        legend.spacing.x = unit(1.5, "lines"), 
        legend.title = element_text(color = "white", hjust = 0.5, size = 25, family = "Segoe UI")) +
       # legend.margin = margin(b = 1, unit = "cm")) +
guides(fill = guide_legend(
    keywidth = unit(2, "lines"),  # Adjust key width
    keyheight = unit(2, "lines"),  # Adjust key height
    label.position = "bottom",  # Position labels below keys
    label.hjust = 0.5,  # Center labels horizontally
    label.vjust = 1,  # Adjust vertical positioning of labels
    nrow = 1, 
    shape = guide_legend(override.aes = list(shape = 19))
  )) 


# save plot
ggsave(paste0(here(), "/output/aqli_intro_map_vis_cap.png"), aqli_global_gadm2_map, width = 17, height = 9, dpi = 320, scale = 2)

  



## Part 1 fig 2: Comparing PM2.5 with other health threats-----------------------------------------------------


# create a version of the figure with the same diseases as used in the same figure in last year's report
 
pm2.5_comparison_oth_threats_data <- gbd_results_master_2021 %>%
  filter(country == "Global", cause_of_death %in% c("PM2.5 relative to WHO guideline", "Tobacco", "Alcohol use", 
                                                    "Unsafe water, sanitation, and handwashing", 
                                                    "Transport injuries", 
                                                    "HIV/AIDS and sexually transmitted infections", 
                                                    "Neglected tropical diseases and malaria", 
                                                    "Nutritional deficiencies", 
                                                    "Child and maternal malnutrition"))
colnames(pm2.5_comparison_oth_threats_data)[3] <- c("llpp_who_2021")

pm2.5_comparison_oth_threats_data <- pm2.5_comparison_oth_threats_data %>%
    mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket))





# pm2.5 compared to other health threats plot

f1 <- "Rubik"

pm2.5_comparison_oth_threats_plt <- pm2.5_comparison_oth_threats_data %>%
  mutate(llpp_who_2021 = round(llpp_who_2021, 1), 
         cause_of_death = ifelse(cause_of_death == "PM2.5 relative to WHO guideline", "PM₂.₅ relative to WHO guideline", cause_of_death)) %>%
  ggplot(mapping = aes(x = forcats::fct_reorder(cause_of_death, llpp_who_2021), y = llpp_who_2021)) + 
  geom_blank() +
  #     annotate("segment", y = c(0.25, 0.75, 1.25), yend = c(0.25, 0.75, 1.25), x = 0, xend = 9,
  # colour = "#0C475F") +
# geom_hline(mapping = aes(yintercept = 0.25), color = "#0C475F", linetype = "solid") +
#   geom_hline(mapping = aes(yintercept = 0.75), color = "#0C475F", linetype = "solid") +
#   geom_hline(mapping = aes(yintercept = 1.25), color = "#0C475F", linetype = "solid") +
  geom_col(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), width = 0.4, color = "white") +
  geom_text(aes(label = llpp_who_2021, y = (llpp_who_2021 - 0.095)), hjust = -0.1, color = "#663300", size = 7, family = f1, 
            fontface = "bold") +
  labs(x = "", y = "Life years lost", fill = "Life years lost", 
       title = expression("How does" ~ PM[2.5] ~ "compare with other heath threats?"), 
       caption = "Sources · Air Quality Life Index, Annual Report 2023; Global Burden of Disease; WHO life tables | Graphic · Aarsh Batra · github.com/AarshBatra | explained.media") +
  coord_flip() + 
  ggthemes::theme_tufte() +
  scale_y_continuous(breaks = seq(0, 3, 0.5), limits = c(0, 2.5)) +
 # scale_x_discrete(limits = cause_of_death_ordered[seq(1, length(cause_of_death_ordered), by = 2)]) +
  scale_fill_manual(values = c("0 - < 0.1" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) + 
   theme(legend.position = "none", 
        legend.background = element_rect(fill = "#022531"), 
        legend.text = element_text(color = "white", face = "bold" , size = 9),
        plot.background = element_rect(colour = "#022531",
                                       fill = "#022531"), 
        plot.caption = element_text(size = 14, hjust = 0, color = "white", margin = margin(t = 1, unit = "cm")), 
        plot.title = element_text(size = 15, family = "Segoe UI", face = "bold", color = "white", hjust = -3.3, 
                                  margin = margin(b = 0, unit = "cm")),
        plot.subtitle = element_text(size = 11, family = "Segoe UI", face = "italic", color = "white", hjust = 0, 
                                     margin = margin(b = 0.5, unit = "cm")),
        legend.spacing.x = unit(1.5, "lines"), 
        legend.title = element_text(color = "white", hjust = 0.5, size = 15, family = "Segoe UI"), 
        axis.text.y = element_text(color = c("PM₂.₅ relative to WHO guideline" = "#FFFFFF", 
                                             "Tobacco" = "#FFE6B3", 
                                             "Child and maternal malnutrition" = "#FFE6B3", 
                                             "Alcohol use" = "#FFE6B3", 
                                             "Unsafe water, sanitation, and handwashing" = "#FFD25D", 
                                             "Transport injuries" = "#FFD25D", 
                                             "HIV/AIDS and sexually transmitted infections" = "#FFBA00", 
                                             "Neglected tropical diseases and malaria" = "#FF9600", 
                                             "Nutritional deficiencies" = "#FF9600"), hjust = 0.5, size = 20, family = f1),
        axis.text.x = element_text(color = "white", hjust = 0.5, size = 20, family = f1),
        axis.title.x = element_text(color = "white", hjust = 0.99, size = 21, family = f1, margin = margin(b = 0.5, t = 0.2, unit = "cm")),
        axis.line = element_line(color = "#0C475F")) +
      ggtext::geom_textbox(aes(x = 4.9,
                             y = 1.3),
                         label = "How does the threat from **<span style=\"color:#CC6633\">particle pollution</span>** compare with other global health threats?",
                         hjust = -0.005, halign = 0,
                         vjust = 1, valign = 1,
                         family = f1,
                         lineheight = 0.35,
                         width = unit(35, "lines"),
                         size = 13,
                         box.colour = NA,
                         fill = NA,
                         color = "white") 
# save plot
ggsave(paste0(here(), "/output/pm2.5_compared_to_other_threats_vis_cap.png"), pm2.5_comparison_oth_threats_plt, width = 10, height = 8, dpi = 320, scale = 2)

ggsave(paste0(here(), "/[upd]pm2.5_compared_to_other_threats_vis_cap.png"), pm2.5_comparison_oth_threats_plt, width = 10, height = 8, dpi = 320, scale = 2)

```



## [TESTING - NOT READY YET] Story other (1): Rural Accesibility Index

```{r}

#### read and clean the data files=====================================================================

#> Define the target projection (Equal Earth - EPSG:8857)
equal_earth_proj <- "+proj=eqearth +datum=WGS84 +wktext"

#> read in country continent mapping, clean col names and convert all character cols to lowercase
country_continent <- read_csv("./data/raw/country_continent.csv")
country_continent$country <- str_to_lower(country_continent$country)
country_continent$continent <- str_to_lower(country_continent$continent)
country_continent <- country_continent %>%
  clean_names()


#> subnational RAI read and clean-------------------------------------------------------
rai_subnational_shp_raw <- st_read("./data/raw/sdgi-9-1-1-rai-2023-shp/sdgi_9_1_1_rai_2023_subnational.shp")
rai_subnational_shp_raw <- st_transform(rai_subnational_shp_raw, equal_earth_proj)
rai_subnational_shp_raw <- rai_subnational_shp_raw %>%
  clean_names() %>%
  mutate_if(is.character, tolower) %>%
  rename(country = name_0, 
         state = name_1, 
         district = name_2) %>%
  select(-c(shape_area, shape_leng))

#> add continent info and add buckets for RAI
rai_subnational_shp_raw_cont <- rai_subnational_shp_raw %>%
  left_join(country_continent, by = "country") %>%
  mutate(continent = ifelse(country %in% c("republic of congo"), "africa", continent)) 

#> add sdg911pct buckets
rai_subnational_shp_raw_cont_sdg_buckets <- rai_subnational_shp_raw_cont %>%
  mutate(sdg911pct_bucket = ifelse(sdg911pct < 0, "Data not available", NA),
         sdg911pct_bucket = ifelse(sdg911pct >= 0 & sdg911pct < 25, "0 to < 25", sdg911pct_bucket), 
         sdg911pct_bucket = ifelse(sdg911pct >= 25 & sdg911pct < 50, "25 to < 50", sdg911pct_bucket), 
         sdg911pct_bucket = ifelse(sdg911pct >= 50 & sdg911pct < 75, "50 to < 75", sdg911pct_bucket), 
         sdg911pct_bucket = ifelse(sdg911pct >= 75, ">= 75", sdg911pct_bucket)) %>%
  mutate(sdg911pct_bucket_order = ifelse(sdg911pct_bucket == "0 to < 25", 1, NA), 
         sdg911pct_bucket_order = ifelse(sdg911pct_bucket == "25 to < 50", 2, sdg911pct_bucket_order), 
         sdg911pct_bucket_order = ifelse(sdg911pct_bucket == "50 to < 75", 3, sdg911pct_bucket_order), 
         sdg911pct_bucket_order = ifelse(sdg911pct_bucket == ">= 75", 4, sdg911pct_bucket_order), 
         sdg911pct_bucket_order = ifelse(sdg911pct_bucket == "Data not available", 5, sdg911pct_bucket_order)) %>%
  mutate(pol_area_sq_met = as.numeric(st_area(geometry))) %>%
  filter(pol_area_sq_met >= 10000000) %>%
  select(-geometry, geometry)


#> national RAI-----------------------------------------------------------
rai_national_shp_raw <- st_read("./data/raw/sdgi-9-1-1-rai-2023-shp/sdgi_9_1_1_rai_2023_national.shp")

#> transform CRS and clean column names, convert everything to lowercase
rai_national_shp_raw <- st_transform(rai_national_shp_raw, equal_earth_proj)
rai_national_shp_raw <- rai_national_shp_raw %>%
  clean_names() %>%
    mutate_if(is.character, tolower) %>%
    rename(country = name_0) %>%
  select(-c(shape_area, shape_leng))

#> add continent information
rai_national_shp_raw_cont <- rai_national_shp_raw %>%
  left_join(country_continent, by = c("country")) %>%
    mutate(pol_area_sq_met = as.numeric(st_area(geometry))) %>%
  filter(pol_area_sq_met >= 10000000) %>%
   select(-geometry, geometry)

#### Plots=========================================================================================


#> plt 1 (ggridges)---------------------------------------

ridges_plt_data_africa <- rai_subnational_shp_raw_cont_sdg_buckets %>%
  st_drop_geometry() %>%
  select(country, state, continent, district, sdg911pct) %>%
  filter(sdg911pct >= 0, continent == "africa")

ridges_plt_data_summary_countries_for_plt <-  ridges_plt_data_africa %>%
  group_by(country) %>%
  summarise(mean_sdg911pct = mean(sdg911pct, na.rm = TRUE)) %>%
  slice_min(mean_sdg911pct, n = 10) %>%
  select(country) %>%
  unlist() %>%
  as.vector()

plt <- ridges_plt_data_africa %>%
  filter(country %in% ridges_plt_data_summary_countries_for_plt) %>%
  ggplot() +
  ggridges::geom_density_ridges_gradient(mapping = aes(x = sdg911pct, y = country, fill = stat(x))) +
  scale_fill_viridis(option = "cividis", direction = -1) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  theme_clean()






#### plt2: plot subnational plot------------------------------------------

gg_record(dir = "tidytuesday-temp2", device = "png", width = 8.44, height = 10.5, units = "in", dpi = 320, scale = 4)

gg_resize_film(
  height = 4,
  width = 4,
  units = "in",
  dpi = 320
)

# # Custom color palette with pomological colors
# palette <- c("#7b8f4e", "#adc178","#dde5b6","#A88469","#8A6950","#6c584c")
# palette <- c("#7b8f4e", "#adc178","#dde5b6","lavenderblush2", "lavenderblush3", "lavenderblush4")
# palette <- c("#6c584c", "#8A6950", "#A88469", "gray75", "gray45", "gray25")
# palette <- c("#7b8f4e", "#adc178","#dde5b6","honeydew4", "honeydew3", "honeydew4")
# palette <- c("#565d50", "#616853", "#6c725b", "#777e62", "#82886a", "#8d9171", "#adc178","#dde5b6")
# palette <-  c("#6b452b", "#a89985", "#efe1c6", "#f8eed1", "#fffeea")
# palette <- c("#e7efc5","#bfd7b5","#a3c4bc","#7B9E91","#3a606e")
# plt1: subnational map-------------------

# Convert sdg911pct to numeric if it's not already
# rai_subnational_shp_raw$sdg911pct <- as.numeric(as.character(rai_subnational_shp_raw$sdg911pct))
# rai_subnational_shp_raw <- rai_subnational_shp_raw %>%
#   mutate(sdg911pct = ifelse(is.na(sdg911pct)))

# read background image
vil_near_highway_img <- png::readPNG("./vil_near_highway_dalle2_6.png")

custom_paragraph <- "<p style='font-family: Arial, sans-serif; font-size: 12px; color: #666666;'>This plot visualizes the Rural Accessibility Index for countries in Africa. The colors represent different accessibility levels, ranging from low to high.</p>"


# plt 1 (map)
rai_subnational_shp_raw_cont_sdg_buckets %>%
  filter(continent == "africa") %>%
 ggplot() +
  # background_image(vil_near_highway_img) +
   geom_sf(mapping = aes(fill = forcats::fct_reorder(sdg911pct_bucket, sdg911pct_bucket_order)), color = "transparent", lwd = 0.02) +
  geom_sf(data = rai_national_shp_raw_cont %>% filter(continent == "africa") %>% select(country, geometry), fill = "transparent", color = "black", lwd = 0.1) +
    geom_sf(data = rai_national_shp_raw_cont %>% filter(continent == "africa", country %in% ridges_plt_data_summary_countries_for_plt) %>% select(country, geometry), fill = "transparent", color = "black", lwd = 0.4) +
  scale_fill_viridis(discrete = TRUE, option = "magma", direction = -1) +
  
  
  # scale_fill_manual(values = c("0 to < 25" = "#c3bd89",
  #                              "25 to < 50" = "#bdae6a",
  #                              "50 to < 75" = "#928877",
  #                              ">= 75" = "#6a7ba0",
  #                              "Data not available" = "lightgrey")) +
 # scale_fill_gradientn(colors = palette, breaks = seq(0, 100, 20), minor_breaks = seq(10, 90, 20)) + 
  # scale_fill_viridis(discrete = TRUE, direction = -1, option = "E") +
  labs(fill = "", title = "Rural Accessibility Index (RAI) in Africa", 
       subtitle = "Proportion of the rural population living within 2 km of an all-season road. <br> <span style='color: #FFEA46FF'><b>Less than 25%</b></span>, <span style='color: #CBBA69FF'><b>25% to 50%</b></span>, <span style='color: #958F78FF'><b>50% to 75%</b></span>, <span style='color: #31446BFF'><b>75% and above</b></span>") +
  theme_map() +
  theme(
    legend.position = "none",  
    plot.background = element_rect(fill = "#aba7cd", color = "#aba7cd"), 
    plot.title = element_markdown(family = "Mono", size = 10), 
    plot.subtitle = element_markdown(family = "Arial", size = 8)
  ) +
 coord_sf(crs = "+proj=eqearth +datum=WGS84 +wktext")

# +
#   guides(fill = guide_legend(
#     keywidth = unit(1, "lines"),  # Adjust key width
#     keyheight = unit(1, "lines"),  # Adjust key height
#     label.position = "bottom",  # Position labels below keys
#     label.hjust = 0.5,  # Center labels horizontally
#     label.vjust = 1,  # Adjust vertical positioning of labels
#     nrow = 1, 
#     shape = guide_legend(override.aes = list(shape = 19))
#   )) 
  

ggsave("./test.png", plt, width = 7, height = 7, units = "in", limitsize = FALSE, dpi = 520)

```


